### 2022-03-17 bools/bools.tex
new bib and paper
wan

Most provers natively support both functional and Boolean extensionality
reasoning: Bhayat et al. \cite{br-20-full-sup-w-combs} modify first-order
unification to return unification constraints consisting of pairs of terms of
functional type, whereas Steen relies on the unification rules of Leo-III's
calculus \cite[Section 4.3.3]{as-18-phd} to deal with extensionality. As a
pragmatic extension of the \lsup{} calculus, Bentkamp et al.
\cite{bbtvw-21-sup-lam} propose to alter the core generating inference rules of
superposition (Sect.~\ref{sec:pre:rules}) to support functional extensionality.
Instead of requiring that terms involved in the inference are unifiable, it is
required that they can be decomposed into \emph{disagreement pairs} such that at
least one of the disagreement pairs is of functional type. Disagreement pairs of
terms $s$ and $t$ of the same type are defined inductively using function
$\textsf{dp}$: $\textsf{dp}(s,t) = \emptyset$ if $s$ and $t$ are equal;
$\textsf{dp}(a \, \tuplen{s}, b \, \tuple{t}{m}) = \{(a \, \tuplen{s}, b \,
\tuple{t}{m})\}$ if $a$ and $b$ are different heads; $\textsf{dp}(\lam{x}{s},
\lam{y}{t}) = \{ (\lam{x}{s}, \lam{y}{t}) \}$; $\textsf{dp}(a \, \tuplen{s}, a
\, \tuplen{t}) = \bigcup_{i=1}^{n} \textsf{dp}(s_i, t_i)$. Then the
extensionality rules are stated as follows:
\pagebreak[2]
\begin{align*}
  & \namedinference{AbsSup}
  {s \eq t \llor C \qquad u[s'] \mathrel{\dot{\eq}} v \llor D}
  {\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor  u[t] \mathrel{\dot{\eq}} v \llor C \llor D}} \\
  &\namedinference{AbsER}{s \noteq s' \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \cdots \llor  C}} \\
  &\namedinference{AbsEF}{s \eq t \llor  s' \eq u \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor t \noteq u \llor  s' \eq u \llor C}}
\end{align*}

@@

\ourpara{Axiomatization}
\looseness=-1
A simple axiomatization of the theory of Booleans is given by Bentkamp et al.\ \cite{bbtvw-21-sup-lam}.
Following their approach, we introduce the proxy type $\typ{bool}$, which corresponds to $o$, to the
signature. We define proxy symbols $\cst{t}, \cst{f}, \cst{not}, \cst{and}, \cst{or},
\cst{impl}, \cst{equiv},\allowbreak \cst{forall}, \cst{exists}, \cst{choice}$, and $\cst{eq}$ which 
correspond to the homologous logical constants from Section \ref{sect:bool:background}. In their
type declarations, $o$ is replaced by $\typ{bool}$.

@@

What our work misses is an overview of heuristics that can be used to curb the
explosion incurred by some of the rules described in this paper. In the next chapter,
we explore exactly this topic.

>>>>

Most provers natively support both functional and Boolean extensionality
reasoning: Bhayat et al. \cite{br-20-full-sup-w-combs} modify first-order
unification to return unification constraints consisting of pairs of terms of
functional type, whereas Steen relies on the unification rules of Leo-III's
calculus \cite[Sect.~4.3.3]{as-18-phd} to deal with extensionality. As a
pragmatic extension of the \lsup{} calculus, Bentkamp et al.
\cite{bbtvw-21-sup-lam} propose to alter the core generating inference rules of
superposition (Sect.~\ref{sec:pre:rules}) to support functional extensionality.
Instead of requiring that terms involved in the inference are unifiable, it is
required that they can be decomposed into \emph{disagreement pairs} such that at
least one of the disagreement pairs is of functional type. Disagreement pairs of
terms $s$ and $t$ of the same type are defined inductively using function
$\textsf{dp}$: $\textsf{dp}(s,t) = \emptyset$ if $s$ and $t$ are equal;
$\textsf{dp}(a \, \tuplen{s}, b \, \tuple{t}{m}) = \{(a \, \tuplen{s}, b \,
\tuple{t}{m})\}$ if $a$ and $b$ are different heads; $\textsf{dp}(\lam{x}{s},
\lam{y}{t}) = \{ (\lam{x}{s}, \lam{y}{t}) \}$; $\textsf{dp}(a \, \tuplen{s}, a
\, \tuplen{t}) = \bigcup_{i=1}^{n} \textsf{dp}(s_i, t_i)$. Then the
extensionality rules are stated as follows:
\pagebreak[2]
\begin{align*}
  & \namedinference{AbsSup}
  {s \eq t \llor C \qquad u[s'] \mathrel{\dot{\eq}} v \llor D}
  {\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor  u[t] \mathrel{\dot{\eq}} v \llor C \llor D}} \\
  &\namedinference{AbsER}{s \noteq s' \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \cdots \llor  C}} \\
  &\namedinference{AbsEF}{s \eq t \llor  s' \eq u \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor t \noteq u \llor  s' \eq u \llor C}}
\end{align*}

@@

\ourpara{Axiomatization}
\looseness=-1
A simple axiomatization of the theory of Booleans is given by Bentkamp et al.\ \cite{bbtvw-21-sup-lam}.
Following their approach, we introduce the proxy type $\typ{bool}$, which corresponds to $o$, to the
signature. We define proxy symbols $\cst{t}, \cst{f}, \cst{not}, \cst{and}, \cst{or},
\cst{impl}, \cst{equiv},\allowbreak \cst{forall}, \cst{exists}, \cst{choice}$, and $\cst{eq}$ which 
correspond to the homologous logical constants from Sect.~\ref{sect:bool:background}. In their
type declarations, $o$ is replaced by $\typ{bool}$.

@@

What our work misses is an overview of heuristics that can be used to curb the
explosion incurred by some of the rules described in this chapter. In the next chapter,
we explore exactly this topic.

### 2022-03-17 ehoh/ehoh.tex
wan

\looseness=-1
Our logic is a variant of the intensional $\lambda$-free Boolean-free
higher-order logic (\lfhol{}) described by Bentkamp et al.\
%, Blanchette, Cruanes, and Waldmann
\cite[Sect.~2]{bbcw-21-lfho}, which could also be called ``applicative
first-order logic.'' In the spirit of FOOL \cite{kotelnikov-16-fool}, we
%% kotelnikov-15-fool
extend the syntax of this logic by making formulas a special case of terms (without adding logical symbols to the signature),
and its semantics by interpreting the Boolean type $o$ as a domain of
cardinality~2. Functional extensionality can be obtained by adding suitable
axioms \cite[Sect.~3.1]{bbcw-21-lfho}.

@@

In this chapter, substitutions $\sigma$ are partial
functions of finite domain from variables to terms, written $\{ x_1 \mapsto s_1,
\ldots, x_m \mapsto s_m \}$, where each $s_i$ has the same type as $x_i$. The
substitution $\sigma[x \mapsto s]$ maps $x$ to $s$ and otherwise coincides with
$\sigma$. Applying $\sigma$ to a variable beyond $\sigma$'s domain is the
identity. We deviated from the view of substitutions in Section \ref{sec:pre:hol}
as it made proofs in Sections \ref{sec:ehoh:unif-match} and \ref{sec:ehoh:indexing}
easier. It is easy to check that both views are equivalent. We also consider unification 
constraints $s \unif t$ as \emph{ordered} pairs.
%Applying a substitution to a term~$t$ applies it homomorphically to $t$'s
%variables.

>>>>

\looseness=-1
Our logic is a variant of the intensional $\lambda$-free Boolean-free
higher-order logic (\lfhol{}) described by Bentkamp et al.\
%, Blanchette, Cruanes, and Waldmann
\cite[Sect.~2]{bbcw-21-lfho}, which could also be called ``applicative
first-order logic.'' In the spirit of FOOL \cite{kotelnikov-16-fool}, we
%% kotelnikov-15-fool
extend the syntax of this logic by making formulas a special case of terms (without adding logical symbols to the signature),
and its semantics by interpreting the Boolean type $o$ as a domain of
cardinality~2. Functional extensionality, the property that two functions are equal if they
always return the same value when given the same arguments, can be obtained by
adding suitable axioms can be obtained by adding suitable axioms
\cite[Sect.~3.1]{bbcw-21-lfho}.

@@

In this chapter, substitutions $\sigma$ are partial
functions of finite domain from variables to terms, written $\{ x_1 \mapsto s_1,
\ldots, x_m \mapsto s_m \}$, where each $s_i$ has the same type as $x_i$. The
substitution $\sigma[x \mapsto s]$ maps $x$ to $s$ and otherwise coincides with
$\sigma$. Applying $\sigma$ to a variable beyond $\sigma$'s domain is the
identity. We deviated from the view of substitutions in Sect.~\ref{sec:pre:hol}
as it made proofs in Sect.~\ref{sec:ehoh:unif-match} and \ref{sec:ehoh:indexing}
easier. It is easy to check that both views are equivalent. We also consider unification 
constraints $s \unif t$ as \emph{ordered} pairs.
%Applying a substitution to a term~$t$ applies it homomorphically to $t$'s
%variables.
